Common lagrangian in introductory courses of classical mechanics depends of $q$
and $\dot{q}$. We refer to them as \emph{first-order lagrangian} because the
highest derivative term is $\dot{q}$, i.e. the first derivative of $q(t)$ with
respect to time. If in the lagrangian appears also $\ddot{q}$ it will be a
\emph{second-order lagrangian} and so on. \\

In the following the mathematical machinery, develop in section \ref{section:
lagrangin and hamiltonian formalism}, will be used to study two apparently
similar lagrangian and introduce the core problem of HD systems.

\subsection{First-order vs Second-order lagrangian}
The simplest non-trivial lagrangians are those in polynomial form like

\begin{equation*}
  L_1(q, \dot{q}) = \frac{\dot{q}^2}{2} - \alpha(q) \qquad
  L_2(q, \ddot{q}) = \frac{\ddot{q}^2}{2} - \alpha(q)
\end{equation*}

where $\alpha$ is a smooth function of $q$ only \cite{Chen13}. No Greek indices
appear on $q^{(i)}$ because we limit ourself to one-dimensional systems. $L_1$
is a first-order lagrangian while $L_2$ it's a second-order one.


\paragraph{first-order lagrangian} In order to get the equation of motion we
used the Euler-Lagrange equation \eqref{eq:euler-lagrange}

\begin{equation} \label{eq: first-order_motion_eq_lagr}
  \frac{\partial L_1}{\partial q} -
  \frac{d}{dt}\frac{\partial L_1}{\partial \dot{q}} +
  \underbrace{
  \frac{d^2}{dt^2}\frac{\partial L_1}{\partial \ddot{q}} - \ldots }_0 = 0
  \qquad \Rightarrow \qquad
  \ddot{q} = - \frac{d\alpha(q)}{dq}
\end{equation}

If $q$ is the position coordinate of a unit mass point, the function $\alpha(q)$
can be interpreted as the potential energy and \eqref{eq:
first-order_motion_eq_lagr} is nothing less than Newton's second law. To get the
hamiltonian description apply the recipe in section \ref{subsection:
hamiltonian_formalism}

\begin{equation*}
  \begin{cases}
    Q := q \\
    P := \frac{\delta L_1}{\delta \dot{q}} = \dot{q}
  \end{cases}
  \qquad
  H_1(q, \dot{q}) := \frac{\delta L_1}{\delta \dot{q}} \dot{q} - L_1(q, \dot{q})
\end{equation*}

\begin{equation} \label{eq: first-order_motion_eq_ham}
  H_1(Q, P) = \frac{P^2}{2} + \alpha(Q) \qquad
  \begin{cases}
    \dot{Q} =   \frac{\partial H}{\partial P} = P \\
    \dot{P} = - \frac{\partial H}{\partial Q} = - \frac{d\alpha(Q)}{dQ}
  \end{cases}
\end{equation}

Since $L_1$ does not explicitly depend on time (only through $q$ and $\dot{q}$),
the hamiltonian $H_1$ can be interpreted as the total energy of the system (sum of
kinetic and potential energy).  Combining the two equation in \eqref{eq:
first-order_motion_eq_ham} we restore \eqref{eq: first-order_motion_eq_lagr}.


\paragraph{second-order lagrangian} Proceeding as before, equation of motion
follow from Euler-Lagrange equation \eqref{eq:euler-lagrange}

\begin{equation} \label{eq: second-order_motion_eq_lagr}
  \frac{\partial L_2}{\partial q} -
  \frac{d}{dt}\frac{\partial L_2}{\partial \dot{q}} +
  \frac{d^2}{dt^2}\frac{\partial L_2}{\partial \ddot{q}} -
  \underbrace{
    \frac{d^3}{dt^3}\frac{\partial L_2}{\partial q^{(3)}} + \ldots
  }_0 = 0
  \quad \Rightarrow \quad
  q^{(4)} = - \frac{d\alpha(q)}{dq}
\end{equation}

Equation \eqref{eq: second-order_motion_eq_lagr} never appear in classical
mechanics, but can it describe some kind of system? The study of the hamiltonian
can give us some clue about the total energy of the system.

\begin{equation*}
  \begin{cases}
    Q_1 := q \\
    P_1 := \frac{\delta L_2}{\delta \dot{q}} = -q^{(3)}
  \end{cases}
  \begin{cases}
    Q_2 := \dot{q} \\
    P_2 := \frac{\delta L_2}{\delta \ddot{q}} = \ddot{q}
  \end{cases}
  \quad
  H_1(q, \dot{q}) :=
    \frac{\delta L_2}{\delta \dot{q}} \dot{q} +
    \frac{\delta L_2}{\delta \ddot{q}} \ddot{q} -
    L_1(q, \dot{q})
\end{equation*}

\begin{equation} \label{eq: second-order_motion_eq_ham}
  H_2(Q_1, Q_2, P_1, P_2) = P_1Q_2 + \frac{P_2^2}{2} + \alpha(Q_1)
\end{equation}

There is a significant difference between the two spectra of $H_1$ and $H_2$:
the first is bounded from below while the latter is not (and both are not
bounded from above).

\subsection{Linear Ostrogradskian instability (Ghosts)}
