The ultimate goal of theoretical physics is the description of all natural
phenomena with a rigorous mathematical formalism. Natural phenomena can be
observe at massive scale (planets, starts, galaxy, black hole, $\dots$),
at minuscule scale (molecules, atoms, particles, $\dots$) and in every day life
with our own naked eyes. Almost everyone agree on the fact that the ``invisible
hand'' that governs all such phenomena is one (the Nature) and follows the
same law independent of the scale. In this regard, our description of the Nature
should be given in a manner that accommodate different scale phenomena in a
unify theory: a \emph{theory of everything} (TOE). \\

Nowadays the theory of \emph{General Relativity} describe the massive scale
phenomena while the tiny ones are describe by \emph{Quantum Fields Theory} and
both are corroborated by numerous experimental evidence. Expert believe that a
TOE is a \emph{superstring theory} of some kind~\cite{Smilga17} but not such a
complete theory is known. Since its born ($\approx 35$ year ago) superstring
theory does not provide us a better understanding of the Nature and, moreover,
it have to deal with the fact that there are no known experiments that can
confirm or reject it. \\

So other roads have be taken in order to formulate a TOE\@. One of them start
from the ansatz that TOE is an ordinary field theory. Even when proposing such a
different theory we have to be able to explain the experimental fact about the
curvature out spacetime. Unlike General Relativity, we keep the space and time
flat and imagine our Universe (3+1)--dimensional as a thin film embedded in a
flat higher-dimensional bulk~\cite{Smilga17}.

\begin{displayquote}
  \emph{Our universe is a (3+1)--dimensional bubble soap in a flat
  higher-dimensional space.}
\end{displayquote}

Such a field theory is formulate in this higher-dimensional bulk and in order to
still renormalizable in the underling Lagrangian appear higher derivative terms.
In light of this we have a good reason to study higher derivative systems.
